### 2022-10-03 src/tex/ms.tex
text changes for referee second round

% Define document class
\documentclass[twocolumn, linenumbers]{aastex631}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\newcommand{\cosmic}{\texttt{COSMIC}}
\newcommand{\legwork}{\texttt{LEGWORK}}
\usepackage{xcolor}
\usepackage{showyourwork}
\newcommand{\snotes}[1]{\textcolor{purple}{#1}}
\newcommand{\kb}[1]{\textcolor{teal}{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\shorttitle{Applying a metallicity-dependent binary fraction to DWD 
formation}
\shortauthors{Thiele et al.}

% Begin!
\begin{document}

% Title
\title{Applying the metallicity-dependent binary fraction to double white 
dwarf formation: \\ Implications for LISA}

% Author list
\correspondingauthor{Sarah Thiele}
\email{sarahgthiele@gmail.com}

\author[0000-0001-7442-6926]{Sarah Thiele}
\affiliation{Department of Physics and Astronomy, University of British 
Columbia, 6224 Agricultural Road, Vancouver, BC, V6T 1Z1, Canada}
\affiliation{Canadian Institute for Theoretical Astrophysics, University
of Toronto, 60 St. George Street, Toronto, Ontario, M5S 1A7,
Canada}
\author[0000-0001-5228-6598]{Katelyn Breivik}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, 
162 Fifth Ave, New York, NY, 10010, USA}
\affiliation{Canadian Institute for Theoretical Astrophysics, University
of Toronto, 60 St. George Street, Toronto, Ontario, M5S 1A7,
Canada}
\author[0000-0003-3939-3297]{Robyn E. Sanderson}
\affiliation{Department of Physics and Astronomy, University of 
Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104, USA}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, 
162 Fifth Ave, New York, NY, 10010, USA}
\author[0000-0002-0296-3826]{Rodrigo Luger}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, 
162 Fifth Ave, New York, NY, 10010, USA}

% Abstract with filler text
\begin{abstract}
Short-period double white dwarf (DWD) binaries will be the most prolific 
source of gravitational waves (GWs) for the Laser Interferometer Space 
Antenna (LISA). DWDs with GW frequencies below $\sim1\,\rm{mHz}$ will be 
the dominant contributor to a stochastic foreground caused by overlapping 
GW signals. Population modeling of Galactic DWDs typically assumes a 
binary fraction of $50\%$ and a log-uniform Zero Age Main Sequence (ZAMS) 
orbital period distribution. However, recent observations have shown that 
the binary fraction of close, solar-type stars exhibits a strong 
anti-correlation with metallicity which modulates the ZAMS orbital period 
distribution below $10^4$ days. In this study we perform the first 
simulation of the Galactic DWD population observable by LISA which 
incorporates an empirically-derived metallicity-dependent binary fraction, 
using the binary population synthesis suite \cosmic\ and a 
metallicity-dependent star formation history. We compare two models: one 
which assumes a metallicity-dependent binary fraction, and one with a 
binary fraction of 50\%. We repeat our analysis for three different 
assumptions for Roche-lobe overflow interactions. We find that while 
metallicity impacts the evolution and intrinsic properties of our 
simulated DWD progenitor binaries, the LISA-resolvable populations of the 
two models remain roughly indistinguishable. However, the size of the 
total Galactic DWD population orbiting in the LISA frequency band is 
reduced by more than half when accounting for a metallicity-dependent 
binary fraction for two of our four variations, which also lowers the 
effective foreground. The LISA population remains unchanged in number for 
two variations, highlighting the sensitivity of the population to binary 
evolution prescriptions.
\end{abstract}

% Keywords
\keywords{Binary stars --- Stellar evolution --- GW astronomy}

\section{Introduction} \label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the 
stars which are born with a binary companion, many will undergo 
interactions which bring the two stars closer together, eventually forming 
a close double white dwarf (DWD). Close DWDs, with orbital periods shorter 
than $\sim2.75\,\rm{hr}$ are the largest source by number of mHz 
gravitational waves (GWs) in the Galaxy 
\citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space 
Antenna (LISA) is expected to resolve at least $10^4$ individual DWD 
binaries in the Milky Way and will also observe GW emission from the 
entire Galactic DWD population through the unresolved foreground created 
by overlapping signals at sub-mHz frequencies 
\citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, 
Korol2017, Lamberts2019, Breivik2020a}. The resolved population will 
enable the study of several important aspects of binary evolution like the 
strength of tides \citep{Valsecchi2012}, the stability of mass transfer in 
DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, 
Sepinsky2014, Kremer2015}, and the separation distribution of close DWDs 
\citep{Korol2021} as well as provide a probe of Galactic structure 
\citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and 
strength of the Galactic DWD foreground can also be used as a tool to 
study the structure of the Milky Way \citep{Benacquista2006, 
Breivik2020b}.

The transition between individually resolved DWDs and the confusion 
limited, or unresolved, DWD foreground is expected to occur near 
$\sim1\,\rm{mHz}$ frequencies \citep[e.g. ][]{Ruiter2010}. In the 
confusion limited regime, more than one binary radiates GWs in each LISA 
frequency bin, thus creating a superposition of signals which are unable 
to be disentangled. There are multiple parameters which shape the Galactic 
DWD foreground. Assuming the evolution of each DWD system is driven solely 
by GW emission, the frequency derivative scales proportionally to 
$f_{\rm{GW}}^{11/3}$, causing a pileup of DWDs at lower frequencies 
\citep[e.g.][]{Breivik2020b}. The spatial density of DWDs in the Galaxy 
which defines the distance to each binary also impacts the foreground 
amplitude, since GW strain scales inversely with distance. Star formation 
history assumptions combine these two effects by assigning ages and 
positions to each DWD in the population which determine the present day 
orbital (and thus GW) frequencies as well as distances to each source.

When viewed strictly as a source of noise, the unresolved Galactic DWD 
foreground is the dominant noise source for LISA in the sub-mHz part of 
LISA's frequency range. This extra noise above the detector noise floor 
affects the detection of all other LISA sources including extreme mass 
ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, 
Moore2017}, merging black holes with masses between 
$10^4$--$10^7\,M_{\odot}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and 
cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, 
Caldwell2019}. For sources which have signals buried by the Galactic DWD 
population, the foreground must be carefully analyzed and subtracted 
\citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number 
of resolved DWDs and the height of the unresolved DWD foreground are a 
direct consequence of the number of DWD progenitors which form and evolve 
over the Milky Way's history.

While the binary fraction remains approximately constant across a large 
metallicity range ($-1.5 \leq$ [Fe/H] $< 0.5$) for wide binaries, close OB 
stars, and the stellar Initial Mass Function (IMF) \citep{Moe2017, 
Moe2019}, the binary fraction for solar-type star systems with orbital 
period $P_{\rm{orb}}\leq 10^4$ days (separation $a \leq 10$ AU) shows a 
strong anti-correlation with metallicity \citep[e.g.][]{Badenes2018, 
Moe2019, Mazzola2020, Price-Whelan2020}. Because close DWDs are the 
remnants of close, solar-type binary stars, this anti-correlation plays an 
important role in the formation, evolution, and characteristics of the DWD 
population that LISA will observe.

To date, population synthesis studies of the Galactic population of close 
DWDs have either assumed a $100\%$ binary fraction or a $50\%$ binary 
fraction, such that for every three stars formed, two reside in a binary 
system \citep{Nelemans2001a, Yu2013, Korol2017, Lamberts2019}. In this 
study, we investigate the effects of a metallicity-dependent binary 
fraction on the formation and evolution of DWDs. To this end, we create 
synthetic present-day Milky Way-like galaxies of DWDs and specifically 
select systems with GW signals that may be observable by the space-based 
detector LISA. Throughout, we make comparisons between the standard 
assumption of a constant $50\%$ initial binary fraction (hereafter model 
F50) and one with a metallicity-dependent binary fraction (hereafter model 
FZ).

In Section \ref{sec:simulations} we discuss our assumptions used in 
simulating DWD populations and detail the process to produce present-day 
synthetic Milky-Way-like galaxies. In Section \ref{sec:LISA_obs} we review 
the derivation of LISA detectability for circular DWD populations at mHz 
frequencies. In Section \ref{sec:results} we detail results showing how a 
metallicity-dependent binary fraction affects the formation and evolution 
of DWD populations assuming four sets of of binary evolution assumptions. 
In Section \ref{sec:LISA_met} we detail the metallicity dependence of the 
LISA DWD foreground and resolved population. Finally, we consider how a 
metallicity-dependent close binary fraction impacts the height of the 
Galactic DWD foreground in LISA for different binary evolution assumptions 
in Section \ref{sec:model_compare} and conclude in Section 
\ref{sec:conclusions}.

\section{Simulating a Galactic DWD population}\label{sec:simulations}
In this section we describe the setup of our DWD simulations using the 
binary population synthesis suite \cosmic, and the process to scale these 
simulations to create Milky Way-like galaxies using the star formation 
history of galaxy \textbf{m12i} in the Latte suite of the FIRE-2 
simulations \citep{Wetzel2016, Hopkins2018} and stellar positions assigned 
according to the the Ananke framework \citep{Sanderson2020}.

\subsection{Metallicity-dependent binary 
fraction}\label{subsec:metbinfrac}
The close binary fraction in the Galaxy has been empirically shown to 
depend on metallicity. This is manifested through an explicit dependence 
of the companion frequency on orbital period \citep{Moe2021}. When 
anchored to a log-normal orbital period distribution, the \emph{frequency} 
of solar-type binary companions is skewed to shorter orbital periods. The 
companion frequency differs from the binary fraction: the companion 
frequency is the probability of a star having a companion at a given 
orbital period, and the integration of this probability over all periods 
results in the binary fraction of the population. The metallicity 
dependence of the companion frequency is observed only below 
$P_{\rm{orb}}=10^6$ days, above which there are not sufficient 
observational data of binary systems to infer a metallicity dependence.

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/met_dep_porb.pdf}
    \caption{The distribution of the companion frequency f$_{\rm{logP}}$ 
for solar-type stars is dependent on both ZAMS orbital period and 
metallicity. Each curve is coloured by metallicity, with the darkest curve 
corresponding to the lowest-metallicity bin in our simulations at [Fe/H] 
$=-2.3$ and the lightest curve corresponding to our highest-metallicity 
bin at [Fe/H] $=0.18$. Above $\log_{10}(P_{\rm{orb}}/\rm{day})\sim 4$, the 
companion frequency distribution is the same for all metallicity bins as 
no identified metallicity dependence at those periods exists in current 
literature.}
    \label{fig:met_dep_porb}
    \script{met_dep_porb.py}
\end{figure}

Figure \ref{fig:met_dep_porb} shows the companion frequency 
f$_{\rm{logP}}$ for solar-type stars as a function of the logarithm of 
orbital period, mirroring the inner binaries curve of Figure 2 in 
\citet{Moe2021} and scaling to an overall multiplicity frequency of 0.67 
(equivalent to an integrated binary fraction of 50\%) at solar 
metallicity. To obtain the normalization of the companion frequency for 
close binaries ($P_{\rm{orb}}<10^4$), we fit the results presented in 
\citet{Moe2019} using linear regression to obtain a piecewise relation 
between the metallicity, [Fe/H] and \textbf{close} binary fraction 
$f_{\rm{b}}$ as

\[f_{\rm{b}}= \begin{cases} 
      - 0.0648 \cdot [\rm{Fe/H}] + 0.3356, & [\rm{Fe/H}]\leq -1.0 
\numberthis \\
      -0.1977 \cdot [\rm{Fe/H}] + 0.2025, & [\rm{Fe/H}]> -1.0  , 
   \end{cases}
\]

\noindent where we convert between [Fe/H] and metallicity Z, assuming all 
stars have solar abundance such that
\begin{equation}
\label{eq:metallicity}
    [\rm{Fe/H}] = 
\log_{10}\Big(\frac{\textit{Z}}{\textit{Z}_{\odot}}\Big),
\end{equation}
\noindent where we assume $Z_{\odot}=0.02$.

We count all higher-order companions as secondary, or binary, components. 
We show curves for all 15 of our metallicity bins, increasing in 
metallicity from darkest to lightest curve colour. The lowest metallicity 
($\rm{[Fe/H]}=-2.3$) shows the highest propensity for short-period 
binaries, in contrast to higher metallicity systems which have a frequency 
distribution with a relatively constant companion frequency for orbital 
periods between $10^4$ and $10^6$ days.

While we simulate binaries across the entire orbital period distribution, 
the close binary fraction, which exhibits the discussed metallicity 
dependence, will be that which impacts the LISA population since systems 
with $\log_{10}(P_{\rm{orb}}/$day)$>4$ will be unlikely to evolve enough 
to enter the LISA band by present day.

\subsection{Binary population models}
\label{sec:bin_pop}
We simulate the evolution of DWD progenitor populations using 
\cosmic\footnote{https://cosmic-popsynth.github.io}, an open-source 
Python-based rapid binary population synthesis suite which employs single 
and binary star evolution using \texttt{SSE/BSE} \citep{Hurley2000, 
Hurley2002}. Several modifications have been added to \cosmic\ which 
incorporate updates for massive star evolution and binary interactions. 
For a detailed description of these modifications see 
\citet{Breivik2020a}. \cosmic\ has been used in several studies to examine 
the effects of binary evolution on binary populations from blue stragglers 
\citep{Leiner2021} and heartbeat stars \citep{Jayasinghe2021}, to white 
dwarf populations \citep{Kremer2017,Breivik2018,Kilic2021}, to merging 
compact object populations in isolated binaries \citep{Zevin2020b, 
Zevin2020a, Zevin2021, Wong2021, Mandhai2021} and in dynamical 
environments around super-massive black holes \citep{Stephan2019, 
Wang2021}.

\cosmic\ is especially useful for efficient generation of large 
populations of compact binaries. Instead of choosing a fixed number of 
binary stars for each simulation, \cosmic\ iteratively simulates 
populations until parameter distributions of the binary population 
converge to a stable shape as more binaries are added. This process is 
quantified through the $match$ parameter inspired by matched filtering 
techniques \citep[e.g. Eq. 6 of ][]{Chatziioannou2017} defined as

\begin{equation}
   match = 
\frac{\sum_{k=1}^{N}P_{k,i}P_{k,i+1}}{\sqrt{\sum_{k=1}^{N}P_{k,i}P_{k,i}\sum_{k=1}^{N}P_{k,i+1}P_{k,i+1}}},
\end{equation}
where $P_{k,i}$ represents the height of bin $k$ on the $i\rm{th}$ 
iteration \citep{Breivik2020a}. In this study, we simulate binaries until 
$\log_{10}(1-match) \leq -5$ for the masses and orbital periods of each 
DWD population at the formation of the second WD. Since all DWD progenitor 
binaries simulated with \cosmic\ are circularized through mass transfer or 
tides before the second WD forms \citep[e.g.][]{Marsh2004, Gokhale2007, 
Sepinsky2014, Kremer2015}, we do not consider convergence of DWD 
eccentricities.

The masses and orbital periods at the formation of the second-formed WD 
span a wide range depending on the WD binary component types, thus we 
consider four DWD combinations: two helium WDs (He + He), a carbon-oxygen 
WD orbiting a helium WD (CO + He), two carbon-oxygen WDs (CO + CO), and an 
oxygen-neon orbiting a helium, carbon-oxygen, or oxygen-neon WD (ONe + X). 
For each DWD type we simulate a grid of $15$ metallicities spaced 
uniformly in $\log_{10}(Z)$ between $Z=10^{-4}$ to $0.03$, to account for 
the limits of the \citet{Hurley2000} stellar evolution tracks employed in 
\cosmic. This results in a total of $60$ populations across all DWD types 
and metallicities for each set of model assumptions. The output of 
\cosmic\ contains information limited to intrinsic binary properties like 
mass and orbital period. External parameters like Galactic position and 
orientation are assigned in a post-processing scheme which uses 
metallicity-dependent positions and ages from the Ananke framework of 
galaxy \textbf{m12i} from the Latte Suite of the FIRE-2 simulations (see 
Section~\ref{sec:FIRE} for details).

We assume that the Zero Age Main Sequence (ZAMS) masses, orbital periods 
and eccentricities for each binary are independently distributed. We 
choose primary masses following \citet{Kroupa2001}, a flat mass ratio 
distribution \citep{Mazeh1992, Goldberg1994} and a uniform eccentricity 
distribution following \cite{Geller2019}. For our simulations which follow 
a metallicity-dependent binary fraction (model FZ), a skewed log-normal 
orbital period distribution is sampled in accordance with the 
metallicity-dependent companion frequency of \citet{Moe2021} as discussed 
in Section \ref{subsec:metbinfrac}. For our simulations which assume a 
constant 50\% binary fraction (model F50), we use a log-uniform period 
distribution following Opik's Law similar to previous studies 
\citep[e.g.][]{Nelemans2001a, Toonen2012, Korol2017, Lamberts2019}. We 
assume a $100\%$ binary fraction in the F50 COSMIC simulations to reduce 
computation time and scale the simulations to a constant $50\%$ binary 
fraction in a post-processing scheme. We initialize all binaries with the 
same evolution time of $13.7\,\rm{Gyr}$ to capture all potential evolution 
within a Hubble time. Although we simulate binaries across the full 
orbital period distribution, we discard DWD binaries which form with 
separations $a\geq 1000\,R_\odot$ during post-processing, since these are 
unlikely to evolve into the LISA band by present day.

We consider a fiducial set of assumptions which follow the \cosmic\ 
defaults described in \citet{Breivik2020a} except for the treatment of 
Roche-lobe overflow (RLO). The stability of RLO mass transfer is 
determined using critical mass ratios resulting from radius-mass exponents 
\citep{Webbink1985, Hurley2002}, where the critical mass ratio is defined 
as the ratio of the donor to accretor mass. We assume critical mass ratios 
following \citet{Claeys2014} which reduce the standard critical mass ratio 
assumptions from \citet{Hurley2002} for main sequence (MS) donors by 
$\sim50\%$ from $3$ to $1.6$ based on the models of \citet{deMink2007} and 
treat WD accretors separately following the models of 
\citet{Soberman1997}. We increase the mass loss rate from the donor 
following Equation 11 of \citet{Claeys2014}. The amount of mass lost 
during RLO from the donor is limited by the overflow factor of the donor 
radius to its Roche radius following \citet{Hurley2002}. The amount of 
mass accepted by the accretor is limited to $10$ times the accretor's mass 
divided by the accretor's thermal timescale. Finally, for RLO mass loss 
which becomes unstable and leads to common envelope evolution (CEE) we 
assume that the donor's binding energy is calculated according to the fits 
detailed in Appendix B of \citet{Claeys2014} and that orbital energy is 
deposited with $100\%$ efficiency into unbinding the common envelope 
($\alpha=1$).

The LISA-detectable DWD populations that result from our simulations may 
vary depending on the assumptions made regarding binary interactions. In 
order to explore the range of results, we complete this study for three 
binary evolution parameter variations on top of our fiducial set of 
assumptions. For each variation, we consider models FZ and F50 as done in 
the fiducial case described above. In variation $q3$, we vary the 
assumption for the critical mass ratios at which a RLO interaction remains 
stable or becomes unstable from our fiducial assumptions. The critical 
mass ratio $q_c$ is increased to $3.0$ and thus allows stable mass 
transfer for more massive RLO donors. In variations $\alpha25$ and 
$\alpha5$, we modify the common envelope ejection efficiency to be either 
much less ($\alpha=0.25$) or more ($\alpha=5$) than in our fiducial 
assumption ($\alpha=1$) to capture the range of possible CE ejection 
efficiencies quoted in the literature  \citep[e.g.][]{Zorotovic2010, 
Fragos2019}. Larger common envelope ejection efficiencies lead to wider 
post-CE separations, while smaller ejection efficiencies either lead to 
closer post-CE separations or stellar mergers where the envelope ejection 
fails. Between all variations, all other prescriptions for binary 
evolution remain identical besides the one varied parameter. \textbf{A 
summary of all models and variations explored in this work is given in 
Table \ref{tbl:models}.} We delve in depth into the fiducial results in 
each section, and then give a brief overview of how the results change for 
each parameter variation.

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|}
       \hline
       model & binary fraction & orbital period distribution \\
       \hline
       FZ & \citet{Moe2019} & \citet{Moe2021}\\
       F50 & 50\% & flat in $\log(P_{\rm{orb}}/\rm{day})$\\
       \hline
       \hline       
       variation & parameter change & binary evolution change \\
       \hline
       fiducial  & none & none \\
       $\alpha25$ & $\alpha=0.25$ & reduced CE efficiency \\
       $\alpha5$  & $\alpha=5$ & increased CE efficiency\\
       $q3$ & $q_c=3$ & increased critical mass ratio \\
       \hline
    \end{tabular}
    \caption{\textbf{Summary of the models and binary evolution assumption 
variation simulations that are explored in this study.}}
    \label{tbl:models}
\end{table}

\subsection{A metallicity-dependent SFH: Convolving with the FIRE-2 
models}
\label{sec:FIRE}
To create Milky-Way-like galaxies which integrate the 
metallicity-dependent binary fraction, we use the metallicity-dependent 
ages and positions of galaxy \textbf{m12i} from the ``Latte" suite of the 
FIRE-2 simulations \citep{Hopkins2015, Wetzel2016, Hopkins2018} to create 
synthetic, Milky-Way-like DWD populations.

The \textbf{m12i} galaxy provides particle mass resolution of 7070 
$M_\odot$ per star particle. Each star particle has an associated 
metallicity, position, and age, which is combined with the output of 
\cosmic\ to assign DWDs to each star particle by matching its metallicity 
to our \cosmic\ metallicity grid. The positions of each DWD are assigned 
using the Ananke framework since multiple DWD binaries can form within a 
single star particle. Specifically, we use an epanechnikov kernel where 
the kernel size is inversely proportional to the local density to assign 
the radial component of spherically symmetric offsets from the center of 
each star particle following \citet{Sanderson2020}.

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/sfh_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close 
solar-type binaries with P $< 10^4$ days (black) plotted against the 
logarithm of $Z/Z_\odot$, with scatter points denoting the location of the 
metallicity grid used in our \cosmic\ simulations. The secondary axis 
shows the amount of mass in stars formed within each metallicity bin from 
galaxy \textbf{m12i} of the ``Latte" suite in the FIRE-2 simulations as a 
red histogram. The amount of stellar mass formed with super-solar 
metallicities dominates the distribution. Note that the primary axis shows 
$f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of 
mass formed in log-scale. The opposing trends of these two distributions 
compete throughout this study.}
    \label{fig:sfh_vs_fb}
    \script{sfh_vs_fb.py}
\end{figure}

The metallicity-dependent close binary fraction for solar-type binaries 
($P_{\rm{orb}} < 10^4$ days), or simply the close binary fraction from 
here on, is shown in black in Figure \ref{fig:sfh_vs_fb} along with the 
mass in star particles from galaxy \textbf{m12i}, shown in red, as a 
function of our metallicity grid. The close binary fraction, $f_{\rm{b}}$, 
drops drastically across metallicity while the mass formed in 
\textbf{m12i} increases significantly. These two opposing trends compete 
throughout this study along with the impact of metallicity on single star 
evolution to form the final numerical distribution of systems in our DWD 
populations.

Since our \cosmic\ simulations assume a binary fraction of $f_{\rm{b}} = 
1$ for model F50, we scale the amount of mass sampled at ZAMS required to 
produce our \cosmic-generated population of DWDs ($M_{\text{b,ZAMS}}$) to 
the proper amount of mass sampled in single and binary stars 
($M_{\text{ZAMS,sim}}$). We do this by sampling single stars and primary 
masses of binary stars from the \cite{Kroupa2001} IMF and sampling 
secondary masses of the binary stars from a uniform mass distribution, 
where the number of binaries is calculated to ensure $f_{\rm{b}}=0.5$. 
From this sample, we obtain the ratio of mass in single stars to the mass 
in binary stars, $R(f_{\rm{b}})$. For model F50, the ratio is a constant 
$R(f_{\rm{b}}) = 0.64$. The total amount of ZAMS mass in single and binary 
stars is then $M_{\text{ZAMS,sim}} =  M_{\text{b,ZAMS}} (1 + 
R(f_{\rm{b}}))$. For model FZ, no scaling is applied since our \cosmic\ 
simulations already have the population-wide binary fraction incorporated, 
and thus $M_{\text{ZAMS,sim}} =  M_{\text{b,ZAMS}}$.

Once we determine the total ZAMS mass required to produce our simulated 
population for a given metallicity, the number of DWDs formed per unit 
solar mass at metallicity $Z_i$ is
\begin{equation}
    n_{\rm{DWD}}(Z_i) = \frac{N_{\rm{DWD, sim}}(Z_i)}{M_{\rm{ZAMS, 
sim}}(Z_i)}.
\end{equation}

\noindent The number of DWD’s per \textbf{m12i} star particle at 
metallicity $Z_i$ is then
\begin{equation}
    N_{\rm{DWD},\star} (Z_i) = n_{\rm{DWD}}(Z_i)\, M_\star,
\end{equation}

\noindent where $M_\star = 7070\,M_\odot$ is the mass per \textbf{m12i} 
star particle. Since $N_{\rm{DWD},\star} (Z_i)$ is not an integer, we 
treat the decimal component as the probability that the star particle 
contains an extra DWD in addition to the integer number. For each star 
particle, we sample with replacement $N_{\rm{DWD},\star}(Z_i)$ DWDs from 
the corresponding simulated \cosmic\ population at that metallicity and 
assign the ZAMS birth time of each DWD to the formation time of the star 
particle. For most DWD types there is more than one DWD binary system 
assigned to each \textbf{m12i} star particle.

If the DWD formation time is less than the age of the star particle, we 
evolve the DWD over the remaining time between its formation and star 
particle age, $t_{\rm{evol}}$, to produce the present-day population. Once 
a DWD is formed, we assume that the binary evolves only due to the 
emission of GWs. Due to tidal effects and mass transfer between their 
progenitor binaries, all DWDs in our simulations are circular, thus 
eccentricity does not need to be considered. The orbital evolution over 
the time $t_{\rm{evol}}$ is then simply defined according to 
\citet{Peters1964} as
\begin{equation}
    a_f = (a_i - 4\beta t_\text{evol})^{1/4},
\end{equation}
where $a_i$ is the DWD separation at formation, and 
\begin{equation}
    \beta = \frac{64G^3}{5c^5} M_1M_2(M_1+M_2)
\end{equation}
is constant throughout DWD evolution \citep{Peters1964}.

We discard any DWDs for which the sum of their ZAMS birth time, given by 
the star particle formation time, and DWD formation time is larger than 
the age of the star particle since the system will not have evolved long 
enough to become a DWD at present. We further discard any DWDs for which 
the lower mass WD overflows it's Roche lobe before present day, because 
the outcomes of these interactions are highly uncertain and their 
treatment is outside the scope of this work \citep[e.g., ][]{Shen2015, 
Kremer2017}. This choice does not significantly affect our results since 
stably accreting WD binaries are unlikely to contribute appreciably to the 
Galactic DWD foreground due to their small mass ratios at frequencies 
below 1 mHz \citep{Breivik2018}. The separation at which the lower mass WD 
overflows its Roche Lobe is defined as

\begin{equation}
    a_{\text{RLO,}\ell} = R_{\ell} \frac{0.6 q_{\ell}^{2/3} + 
\ln{(1+q_{\ell}^{1/3})}}{0.49 q_{\ell}^{2/3}}
\end{equation}
where $R_{\ell}$ is the radius of the lower mass WD and $q_{\ell} = 
M_{\ell}/M_{h}$ is the ratio of the lower- to higher-mass WD components. 
\citep{Eggleton1983}. 
We define the radius of a WD following \citet{Tout1997, Hurley2000} as
\begin{equation}
    R_{\rm{WD}} = \max\Bigg(R_\text{NS}, 
0.0115\sqrt{\left(\frac{M_\text{Ch}}{M}\right)^{2/3}-\left(\frac{M}{M_\text{Ch}}\right)^{2/3}} 
\Bigg) 
\end{equation}
\noindent where $R_\text{NS} = 1.4\cdot 10^{-5}~R_\odot$ is the radius of 
a neutron star, $M_{\text{Ch}}=1.44 ~M_\odot$ is the Chandrasekhar limit 
for the mass of a stable WD, and $M$ is the mass of the WD in solar 
masses.

For the non-discarded systems, we log the present-day separations from 
which the present-day orbital frequency $f_\text{orb}$ can be found using 
Kepler's third law. The GW frequency is then $f_\text{GW} = 
2f_\text{orb}$.

\section{LISA detectability}
\label{sec:LISA_obs}
We use \legwork\footnote{https://legwork.readthedocs.io} \citep{Wagg2021} 
to determine the detectability of our simulated DWD populations for 
sources with GW frequencies $f_{\rm{GW}}>10^{-4}\,\rm{Hz}$. \legwork\ 
calculates the position-, orientation-, and angle-averaged signal to noise 
ratio (SNR) for inspiraling GW sources closely following the derivations 
of \citet{Flanagan1998} and using the LISA noise power spectral density 
(PSD) of \citet{Robson2019}.

To lowest order in the post-Newtonian expansion, the frequency evolution 
of circular orbits for quadrupole GW emission is defined as
\begin{equation}
    \Dot{f}_n = \frac{48n}{5\pi} \frac{(G\mathcal{M}_c)^{5/3}}{c^5} (2\pi 
f_\text{orb})^{11/3}.
\end{equation}
We classify DWDs as evolving, or ``chirping", when $\Dot{f}_n \geq 
1/t_\text{obs}^2$. For evolving sources, the SNR is

\begin{equation}
   \langle \rho \rangle^2_{\text{circ,evol}} = \int_{f_0}^{f_1}df   
\frac{h_{c}^2}{f^2S_n(f)}
\end{equation}

\noindent where $h_c$ is the characteristic strain of the system, $S_n(f)$ 
is the LISA sensitivity curve of \citet{Robson2019}, and the frequency 
limits are determined by the orbital evolution over the observation time, 
$T_{\rm{obs}}=4\,\rm{yr}$. The characteristic strain for circular orbits 
is 
\begin{equation}
    h_{c}^2 = \frac{2^{2/3}}{3 \pi^{4/3}} \frac{(G 
\mathcal{M}_c)^{5/3}}{c^3 D_L^2} \frac{1}{f_{\rm orb}^{1/3}},
\end{equation}
\noindent where $\mathcal{M}_c = (M_1M_2)^{3/5}/(M_1 + M_2)^{1/5}$ is the 
system's chirp mass, and $D_L$ is the systems luminosity distance which we 
assume to be the distance of each simulated DWD to the Sun.

For stationary sources, the SNR is modified due to the lack of observable 
orbital evolution as
\begin{equation}
    \rho_{\text{circ,stat}}^2 = \frac{h_{2}^2T_{\text{obs}}}{ S_n(f_2)}
\end{equation}
with the observation time $T_{\text{obs}}$. Here, $h_2$ is strain 
amplitude of the source for the second orbital frequency harmonic,
\begin{equation}
   h_2^2 = \frac{2^{22/3}}{5}\frac{(G\mathcal{M}_c)^{10/3}}{c^8D_L^2}(\pi 
f_\text{orb})^{4/3}
\end{equation}
and is connected to the characteristic strain as
\begin{equation}
    h_2^2 = \frac{\dot{f}_{2}}{f_{\rm{orb}}^2} h_c^2
\end{equation}

The amplitude spectral density for a stationary system is finally defined 
as $ASD = h_{2} \sqrt{T_\text{obs}}$, such that the SNR for stationary 
source is simply, $\rho \sim ASD/S_n$.

The Galactic foreground included in the \citet{Robson2019} LISA noise 
curve was generated using a different binary evolution code and set of 
model assumptions for DWD formation and evolution \citep{Toonen2012, 
Korol2017}. Thus, we use the detector curve only and generate an 
approximate foreground from each of our populations as follows. Instead of 
performing a full source subtraction algorithm 
\citep[e.g.][]{Littenberg2020}, which is out of the scope of this work, we 
calculate the PSD of the Galactic DWD population with a frequency 
resolution set by the LISA mission time as 
$1/T_{\rm{obs}}\sim1/4\,\rm{yr}^{-1}\sim8\times10^{-9}\,\rm{Hz}$. We then 
approximate the foreground as the running median of the PSD with a boxcar 
window with a width of $10^3$ frequency bins similar to 
\citet{Benacquista2006}. The Galactic DWD PSD is truncated near 
$10\,\rm{mHz}$ for both of our models because we remove all DWDs which 
experience Roche-lobe overflow. In order to smooth the effect of this 
truncation in our foreground, we fit each running median with fourth-order 
polynomials for GW frequencies up to $1\,\rm{mHz}$, thus allowing an 
approximation of the foreground PSD for higher frequencies. These fits are 
listed in Table~\ref{tbl:fits} where the polynomial is described as
\begin{equation}
\label{eq:fit}
    \log_{10}(\rm{confusion\ fit}/\rm{Hz}) = a\,x^4 + b\,x^3 + c\,x^2 + 
d\,x + e
\end{equation}
\noindent and $\rm{x}=\log_{10}(f_{\rm{GW}}/\rm{Hz})$. We add the fitted 
polynomial of the PSD's running median to the LISA noise PSD to obtain a 
sensitivity curve for each model and variation.

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|c|c|c|}
       \hline
       model & a & b & c & d & e \\
       \hline
       \hline
       fiducial, F50 & -217.8 & -183.9 & -74.5 & -13.7 & -1.0 \\
       \hline
       fiducial, FZ & -268.2 & -243.0 & -100.5 & -18.7 & -1.3 \\
       \hline
       $\alpha25$, F50 & -5520.2 & -6078.3 & -2528.8 & -467.1 & -32.3 \\
       \hline
       $\alpha25$, FZ & -2816.0 & -3066.4 & -1272.7 & -234.7 & -16.2 \\
       \hline
       $\alpha5$, F50 & -178.7 & -142.8 & -58.7 & -11.0 & -0.8 \\
       \hline
       $\alpha5$, FZ & -265.9 & -239.8 & -99.1 & -18.5 & -1.3 \\
       \hline
       q3, F50 & -114.5 & -73.5 & -30.9 & -6.1 & -0.5 \\
       \hline
       q3, FZ & -336.9 & -323.5 & -136.2 & -25.8 & -1.8 \\
       \hline
    \end{tabular}
    \caption{Polynomial fitting coefficients for the confusion foreground 
fit of Equation~\ref{eq:fit} for each binary fraction model and parameter 
variation.}
    \label{tbl:fits}
\end{table}

\section{Metallicity effects on the formation and evolution of 
DWDs}\label{sec:results}

\subsection{DWD types and their formation channels}\label{sec:ini}
As discussed in Section\,\ref{sec:bin_pop}, we consider four DWD 
sub-types, which each contribute differently to LISA's GW signals: He + 
He, CO + He, CO + CO, and ONe + X. Each sub-type has a unique distribution 
in their formation times, initial masses, radii, and orbital periods 
stemming from variations in their evolution channels and their formation 
efficiency. Here we describe the general formation scenarios and 
population properties of Galactic close DWDs which may be observable by 
LISA.

He WDs are unable to form through single star evolution within the 
lifetime of the Milky Way. Instead, they originate through interactions in 
close binary systems or binaries with large eccentricities. Because of 
this, He WDs are able to form with low component masses on order $\sim 
0.1\,M_\odot$, with the majority of He WDs in our simulations having 
masses between 0.2-0.5 $M_\odot$. He + He DWDs form through evolution of 
close binary systems, during which their envelopes are both stripped 
through RLO and CE phase interactions before Helium ignition occurs. The 
two progenitor stars generally have masses $\lesssim 3\,M_\odot$ which is 
lower than the progenitors of other DWD types. Our simulated He + He DWDs 
have an approximately constant distribution of formation times $\gtrsim 
2.5\,\rm{Gyr}$. Lastly, since the ZAMS separations are skewed towards 
shorter values, we also see that the resulting DWD separations are smaller 
on average than that of other DWD types.

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/form_eff.pdf}
    \caption{The DWD formation efficiency vs metallicity of DWD 
populations simulated with \cosmic. Each panel shows the formation 
efficiency for a given DWD type and variation. The solid lines indicate 
the formation efficiency for model FZ which incorporates a 
metallicity-dependent binary fraction. The dashed lines indicate the 
formation efficiency for model F50, which assumes a constant binary 
fraction of $50\%$. The DWD formation efficiency drops by a factor of 
$4$--$5$ for model FZ and a factor of $1$--$5$ for model F50. See 
Section~\ref{sec:formeff} for a careful description of the trends for each 
DWD type.}
    \label{fig:form_eff}
    \script{form_eff.py}
\end{figure*}

A CO WD forms when a star is able to begin the helium burning process 
before its envelope is stripped. Thus to form a CO + He DWD, RLO and CE 
stages occur after one component experiences core helium burning, but 
before the other component can. Most close CO + He DWDs form in 
approximately 2 Gyr after ZAMS and with very short periods because the He 
WD is formed through the ejection of a common envelope which greatly 
reduces the orbital separation. Because of these short formation 
separations, many CO + He DWDs merge before the present day. Due to their 
asymmetric mass distributions, they have lower chirp masses, but their 
shorter periods make them important candidates for LISA detection.

To prevent the two stars' envelopes from being stripped before helium 
ignition, CO + CO DWDs typically form from progenitors in wider orbits, 
and the two components may have little to no interaction during their 
evolution from ZAMS to DWD. CO DWDs thus have a distribution in progenitor 
separation that extends to larger values than for other DWD types. Most CO 
+ CO DWDs need $> 0.3$ Gyr to form, have component WD masses between 
$0.35$--$1.0\,M_\odot$, and make up the majority of the DWD population.

ONe WDs are rare and typically form from massive progenitor stars which 
evolve through the asymptotic giant branch phase, thus resulting in a 
higher mass WD. All ONe WDs in our \cosmic\ populations, e.g., have 
progenitor ZAMS masses above $4\,M_\odot$, and the resulting ONe WDs have 
a relatively flat distribution of masses from $1.05\,M_\odot$ up to the 
Chandrasekkhar limit of $1.4\,M_\odot$. Because an ONe WD can have a 
companion of any other WD type in our study, there is a spread in their 
distributions for separation, secondary mass, final orbital period, and 
formation time. In general, however, these systems result from wider 
separations to allow for the evolution of the ONe component without 
merging. For example, all initial separations in our \cosmic \ populations 
have separations $\gtrsim 1.5\,R_\odot$. ONe + X DWDs can form on short 
timescales, as low as $30\,\rm{Myr}$, for the majority of high-metallicity 
systems.

\subsection{Metallicity-dependent trends in the formation efficiency of 
DWDs}\label{sec:formeff}
The number of DWDs formed per unit solar mass of ZAMS star formation, or 
DWD formation efficiency $\eta_{\rm{form}}(Z)$, varies with metallicity. 
Consequently, a metallicity-dependent binary fraction further impacts the 
efficiency of DWD formation within the Galaxy. Figure \ref{fig:form_eff} 
shows the DWD formation efficiency as a function of metallicity for each 
DWD type, binary fraction model, and binary evolution parameter variation. 
In general, the formation efficiency decreases with increasing 
metallicity. This effect is exaggerated for model FZ which assumes a close 
binary fraction which also decreases with increasing metallicity. In 
\textbf{this section}, we \textbf{give a brief overview of the general 
trends observed in formation efficiency and their underlying causes. For a 
detailed description for each DWD type and binary evolution parameter 
variation, see Appendix \ref{appendix:form_eff}.}

\textbf{There are two common ways that the formation of DWDs from stellar 
progenitors is inhibited. The first is through stellar mergers, where the 
binary merges before both components become WDs. The second is due to 
evolutionary outcomes like forming a different DWD type, or not forming a 
DWD by the present day. For all DWD types except ONe + X, which has other 
dominating affects hindering formation efficiency like 
metallicity-dependent stellar winds, we find that different regimes of 
ZAMS orbital period leads to different specific evolutionary channels.}

\textbf{In our \cosmic\ models, lower-metallicity stars evolve faster than 
high-metallicity stars and have larger radii near the end of the MS. This 
changes the timescales of RLO: low-metallicity stars tend to fill their 
Roche lobes while still on the MS, meaning mass transfer can remain stable 
which serves to widen the binary. In contrast, high-metallicity systems 
initiate mass transfer after the donor has left the main sequence, so the 
binary is more likely to enter a CE phase. Moreover, high-metallicity 
progenitors tend to have deeper convective envelopes such that a larger 
fraction of their mass is contained in the envelope relative to 
lower-metallicity counterparts of the same mass \citep{Amard2019, 
Amard2020}. When these systems enter a CE phase, it thus requires more 
orbital energy to eject the CE than for a low-metallicity star at the same 
point in its evolution, causing a large amount of orbital shrinking. These 
effects often leads to a stellar merger at higher metallicities for a 
system that would have survived the CE phase at a lower metallicity, thus 
causing a decrease in formation efficiency with metallicity.}

The skews introduced by the metallicity-dependent ZAMS orbital period 
distribution and binary fraction of FZ also contribute to the formation 
efficiency trends. The log-normal metallicity-dependent $P_{\rm{orb}}$ 
distribution increasingly skews towards orbital periods below 10$^4$ days 
for decreasing metallicity. This amplifies the effects of 
metallicity-dependent stellar evolution discussed above. For example, the 
skew induces more mergers at high metallicities where the deeper 
convective envelope leads to increased orbital shrinkage. However, it is 
also of note that the log-uniform orbital period distribution of F50 
produces a higher number of the shortest-orbital period systems (below 
$\sim100$ days) than FZ, since the tails of the FZ distribution fall 
quickly below this threshold. These systems also have a propensity for 
merging before the formation of a DWD.

\textbf{Our binary evolution parameter variations emphasizes the interplay 
of the aforementioned effects of the distributions of F50 and FZ with the 
physics of binary evolution. For example, the lower CE efficiency of 
variation $\alpha25$ is the dominating formation efficiency inhibitor 
since it induces a higher merger rate. In contrast, variations $\alpha5$ 
and $q3$ relax hindrances to DWD formation (through an increased ability 
to survive a CE phase and undergo stable mass transfer respectively). We 
thus find less mergers overall in both cases, and that the effect of the 
metallicity-dependent $P_{\rm{orb}}$ distribution dominates the formation 
efficiency trends rather than restrictions imposed by binary evolution 
prescriptions.
}

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/CEsep.pdf}
    \caption{Average interaction separation, 
$\overline{\mathrm{a}}_{\rm{CE}}$ of progenitors of close DWDs across 
metallicity for each DWD type and variation from model F50. Solid lines 
show the average separation at the first RLO for binaries in each 
metallicity bin. The shaded regions show the $1\sigma$ spread around the 
mean within each metallicity bin. The average interaction separation 
increases with metallicity for every DWD type, except for $\alpha5$ which 
remains approximately constant for CO + He and ONe + X. The positive trend 
in the average interaction separation is a direct consequence of larger 
envelope masses of higher-metallicity donors which are less evolved than 
their lower-metallicity counterparts. The sharp drop of the CO + CO 
average at the highest-metallicity bin of the fiducial variation is due to 
binning effects of the figure: it only contains 506 systems, of which the 
majority stem from a single low-separation system which was sampled 468 
times times during the simulation. The regions of metallicity where there 
is not shading beyond the mean trend are also due to low numbers of DWDs 
in each bin.} 
    \label{fig:CEsep}
    \script{CEsep.py}
\end{figure*}

\subsection{Metallicity trends in DWD progenitor common envelope 
separation}\label{sec:CEsep}

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/lisa_nums.pdf}
    \caption{The number of LISA-band systems formed for each DWD type and 
parameter variation as a function of the base-10 logarithm of metallicity, 
normalized to solar value. The solid line shows the FZ population with a 
metallicity-dependent close binary fraction incorporated, and the dashed 
line shows the F50 population for a standard binary fraction of 0.5. The 
LISA population of DWDs is dominated by stars with super-solar 
metallicities. This is true even for model FZ, which drops off 
significantly for higher metallicities, because of the large number of 
stars formed in \textbf{m12i} beyond $Z\simeq Z_\odot$. There is a double 
peak in the fiducial and $\alpha25$ He + He populations; the first peak is 
caused by the sharp drop in formation efficiency past 
$Z\simeq0.1\,Z_\odot$ which is then greatly overcompensated for by the 
amount of star formation at higher metallicities which forms the second 
peak.}
    \label{fig:lisa_nums}
    \script{lisa_nums.py}
\end{figure*}

All systems that end up radiating GWs in the LISA band have undergone at 
least one phase of CE evolution. For systems which experience a stable RLO 
mass transfer in the first interaction, the CE phase plays a key role in 
shrinking systems with initially wide separations to bring them into the 
LISA band.

Figure \ref{fig:CEsep} shows the average separation at the first instance 
of CE evolution, $a_{\text{CE}}$ of all DWD progenitors that result in 
systems orbiting in the LISA frequency band at present day, as a function 
of metallicity for each DWD type. The solid lines denote the average 
value, and the $1\sigma$ variance is shown in the surrounding shading. We 
show this for our fiducial variation and each of our binary parameter 
variations. For our fiducial, $\alpha25$ and $\alpha5$ parameter 
variations, the average CE separation increases in general across 
metallicity. Higher-metallicity binaries will interact earlier in the 
binary's lifetime than a lower-metallicity binary of equal separation due 
to their relatively larger maximum radii. Since higher-metallicity 
binaries are also more likely to merge during CE interactions because of 
their relatively more massive donor envelopes, the DWDs which survive and 
eventually orbit in the LISA band originate from systems with higher 
interaction separations which allow their orbits to shrink significantly 
during CE phases without merging. This regulation plays a key role in 
smearing out any observable effects of a metallicity-dependent binary 
fraction in the population of DWDs observable by LISA.

The slope of the increase for these three models is dictated by the CE 
efficiency $\alpha$. \textbf{As discussed in Section \ref{sec:formeff}, 
with} increasing metallicity there is a deepening of the convective 
envelope of the star \textbf{compared to its lower metallicity 
counterpart}, requiring more orbital energy to be available in order to 
eject the CE. Thus the high-metallicity binaries must have even wider 
separations for lower $\alpha$ values to ensure the binary does not merge 
before the envelope is ejected. We therefore see a steeper separation 
slope for $\alpha25$, a less steep slope for $\alpha5$, and an 
intermediate slope for our fiducial variation which has $\alpha=1$.

The DWDs in model $q3$ that survive into the LISA band at present day have 
different formation channels than those of our other three parameter 
variations. Most importantly, in contrast to two CE phases for the 
fiducial, $\alpha25$ and $\alpha5$ parameter variations, the DWD 
progenitors in parameter variation $q3$ typically undergo a phase of 
stable mass transfer as the binary's first interaction. During mass 
transfer the orbit first shrinks, and then will slightly widen once the 
binary's mass ratio flips. This process leads to formation of the first WD 
component. The companion is then formed later in the binary's evolution 
through a CE phase. We show the CE phase which results from the second 
interaction for parameter variation $q3$ in Figure \ref{fig:CEsep}.

Of particular interest in variation $q3$ is the CO + He systems. For all 
of the $q3$ CO + He binaries that reside in the LISA band at present, the 
He WD unexpectedly forms first through stable mass transfer, with the CO 
WD forming second through a CE. These binaries tend to all have similar 
resulting orbital separations from mass transfer, as can be seen by the 
relatively constant curve for CO + He separation at CE onset. There is 
still a slight increasing trend in CO + He (as well as CO + CO and a more 
pronounced trend in He + He) for $q3$ that stems from the 
metallicity-dependent convective envelope discussion above. We also see 
that the average interaction separation are overall lower for variations 
$q3$ and $\alpha5$ than for the other two due to the orbital shrinkage 
that has already occurred from stable mass transfer and the efficient CE 
phase that will lead to minimal orbital shrinkage respectively.

\section{Metallicity Dependence of the LISA DWD Population}
\label{sec:LISA_met}

While metallicity impacts the intrinsic properties of our simulated DWD 
populations as described in Sections~\ref{sec:formeff} and 
\ref{sec:CEsep}, when we consider the present-day Galactic close DWDs we 
find that the population detectable by LISA largely only changes 
\textbf{in total number, but not in frequency distribution}. The number of 
DWDs in the LISA frequency band decreases by $\sim50\%$ when comparing 
model F50 to model FZ for variations $\alpha5$ and $q3$, but remains 
approximately unchanged for our fiducial and $\alpha25$ variations.

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/PSD.pdf}
    \caption{The PSD of the entire Galactic DWD GW foreground, summed over 
all metallicities and DWD types vs GW frequency for models FZ and F50, 
where the PSD is downsampled by a factor of $10$ for plotting purposes. 
Each panel shows a different parameter variation. The vertical lines show 
the PSD where model F50 is shown in dark blue and model FZ is shown in 
light blue. The dashed lines show fits to the rolling boxcar median of 
width 1000 bins for each PSD (see Section \ref{sec:LISA_obs} for a 
discussion). A metallicity-dependent binary fraction (model FZ) yields 
fewer DWDs across all frequencies than a $50\%$ binary fraction (model 
F50) by roughly a factor of $2$ for the $\alpha5$ and $q3$ parameter 
variations. This produces a lower GW confusion foreground for frequencies 
with f$_{\rm{GW}}\leq\sim10^{-3}$ Hz. In contrast, the fiducial and 
$\alpha25$ parameter variations show very similar confusion foregrounds.}
    \label{fig:PSD}
    \script{PSD.py}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{figures/LISA_SNR_fiducial.pdf}
    \caption{The ASD vs GW frequency for DWDs resolved with SNR $> 7$ for 
each DWD type where the top row shows the population from model F50 and 
the bottom row shows the population from model FZ for the fiducial 
variation. In each panel, the LISA sensitivity curve, including the 
confusion foreground for each model, is shown in black and the total 
population for each model is shown in grey. We find that each model 
qualitatively exhibits similar characteristics and that the only change is 
in the yield of resolved DWDs for each type based on the strength of the 
confusion foreground.}
    \label{fig:LISA_SNR}
    \script{LISA_SNR.py}
\end{figure*}

Figure \ref{fig:lisa_nums} shows the number of DWDs orbiting with 
frequencies $f_{\rm{GW}} > 0.1\,\rm{mHz}$ against metallicity for each DWD 
type and parameter variation. The solid lines show DWDs from model FZ, and 
the dashed lines show DWDs from model F50. The He + CO, CO + CO, and ONe + 
X populations each have strong peaks in the number of DWDs near solar 
metallicity at which the majority of star formation in galaxy {\bf{m12i}} 
occurs. The largest contribution to the population comes from 
metallicities above $\sim 0.01\,Z_\odot$. Any discrepancy between the two 
binary fraction models is also the most significant above this threshold. 
When creating our DWD populations the DWD formation efficiency, number of 
\textbf{m12i} star particles, and the metallicity-dependent binary 
fraction and ZAMS orbital period distribution all compete. The amount of 
stars formed in \textbf{m12i} at higher metallicity values overwhelms the 
drop in DWD formation efficiency by multiple orders of magnitude, so this 
effect dominates when determining the number of stars initially sampled in 
the population for $f_{\rm{b}}$.

There are two peaks in the distribution of He + He DWDs which are most 
prominent in the fiducial and $\alpha25$ variations, and not as obvious 
for $\alpha5$ or $q3$. The prominent first peak in fiducial and $\alpha25$ 
occurs because near $Z\simeq 0.1\,Z_\odot$, the DWD formation efficiency 
transitions from near constant values to a sharp decrease (see Figure 
\ref{fig:form_eff}). However, for a drop in the formation efficiency by a 
factor of of $\sim$ six, the amount of star formation in galaxy 
\textbf{m12i} increases by more than an order of magnitude for supersolar 
metallicities. Above $Z=Z_\odot$ this overcompensates for the efficiency 
drop, producing the second peak. For $\alpha5$ and $q3$, the lack of 
dramatic orbital shrinking -- from the increased CE efficiency in the 
former and increased propensity for stable mass transfer in the latter -- 
allow for less failed DWDs due to stellar mergers. This leads to less 
dramatic drops in formation efficiency across metallicity which thus 
suppresses the formation of two peaks in the number of LISA-band systems.

The $\alpha5$ and $q3$ variations produce a higher number of F50 systems 
than FZ when summed over all DWD types. This is due to the log-uniform 
P$_{\rm{orb}}$ distribution of model F50 having more of the 
shortest-period binaries than the tails of the log-normal distribution of 
FZ; a discrepancy which becomes more significant with increasing 
metallicity as the FZ distribution moves from a peak of 
log$_{10}$($P_{\rm{orb}}\rm{/day})\simeq 3-4$ to 
log$_{10}$($P_{\rm{orb}}\rm{/day})\simeq 4-6$. This tail is less likely to 
survive to the LISA band at present for the fiducial and $\alpha25$ 
variations due to the increased propensity for mergers of DWD progenitors. 
This effect produces an equal number of systems between FZ and F50 for the 
fiducial and $\alpha25$ variations. Although the close binary fraction 
leads to a greater number of low-metallicity systems, this is compensated 
for by the FIRE star particle metallicity distribution as well as the 
tendency for the shortest-period DWD progenitors in model F50 to merge, 
essentially washing out all differences between the populations.

\begin{figure*}

\includegraphics[width=\textwidth]{figures/Mc_vs_dist_fiducial.pdf}
    \caption{The chirp mass vs distance for each DWD type is shown for our 
fiducial parameter variation. Only systems with observable evolution in 
their GW frequency, i.e those which are chirping, and with SNR $>7$, are 
shown, since these are systems for which distance can be separated from 
chirp mass within their strain amplitude. Each panel shows one DWD type, 
summed over all metallicities. Model FZ is indicated with solid light blue 
contours, and model F50 is indicated with dark blue dashed contours 
respectively. Contours are shown at the $5^{\rm{th}}$, 25$^{\rm{th}}$, 
50$^{\rm{th}}$, 75$^{\rm{th}}$, and 95$^{\rm{th}}$ percentiles. Despite 
intrinsic changes to population properties induced by a 
metallicity-dependent binary fraction, and a reduction in the height of 
the DWD Galactic foreground the distributions are very similar.}
    \label{fig:Mc_vs_dist}
    \script{Mc_vs_dist.py}
\end{figure*}

A change in the number of DWDs in model FZ relative to model F50 is also 
apparent in the GW PSD LISA will observe. We show the GW PSD of each 
model, as well as the confusion forground estimate, in 
Figure~\ref{fig:PSD}. We downsample the PSD frequencies by a factor of 10 
for plotting purposes. While model F50 (dark blue) and model FZ (light 
blue) produce several thousand large spikes in the PSD across LISA's 
frequency band, the overall foreground height, including the confusion, is 
larger for model F50 for parameter variations $\alpha5$ and $q3$. This is 
a direct consequence of the overall reduction in the size of the close DWD 
population in model FZ. The foreground height remains unchanged for the 
fiducial and $\alpha25$ variations when comparing models FZ and F50.

Similar to previous studies, we find that LISA will be able to resolve 
several thousand DWDs. Figure \ref{fig:LISA_SNR} shows the amplitude 
spectral density vs GW frequency of the resolved systems with SNR $> 7$ 
for each DWD type, where the top and bottom rows show results for models 
FZ and F50 respectively for the fiducial parameter variation. For 
comparison, we also show the LISA sensitivity curve, including the 
modeled confusion foreground from each population in black, and the the 
entire population for each model in grey. Plots for our three parameter 
variations can be found in Appendix \ref{appendix:ASD_vars}. Apart from 
each DWD type having a different abundance of resolved systems, the 
population-wide characteristics remain unchanged between the two binary 
fraction models. The populations containing at least one He WD occupy the 
lower-ASD, higher-GW frequency region of parameter space compared to the 
total population, with CO + He DWDs having larger ASDs than the He + He 
DWD population. Conversely, DWD types without a He WD component tend to 
occupy the higher-ASD, lower-GW frequency region of parameter space. This 
difference is largely due to the formation scenario of DWDs containing a 
He WD, which form from the ejection of a common envelope created by the He 
WD progenitor. These lower mass progenitors overflow their Roche lobes at 
closer separations relative to the higher mass progenitors (e.g. 
Figure~\ref{fig:CEsep}) and thus also produce closer DWDs. While the 
distance to any one DWD strongly influences its ASD, DWD populations 
without a He WD component have, on average, higher ASDs due to their more 
massive WD components.

We note that the height of the confusion foreground and the number of DWDs 
radiating GWs with $f_{\rm{GW}} > 10^{-4}\,\rm{Hz}$ is reduced by a factor 
of 2 for model FZ relative to model F50 for variations $\alpha5$ and $q3$, 
the number of resolved sources is not reduced to an equal degree. This is 
because in the absence of less competing GW signals from DWDs in the LISA 
frequency band, more DWDs can be individually resolved. For the fiducial 
and $\alpha25$ variations, the number of resolvable systems with SNR $>7$ 
are reduced by 4.5\% and increased by 3\% respectively. However, for 
$\alpha5$ the number of resolvable DWDs is reduced by only 26\%, and the 
number of resolvable DWDs $q3$ remains almost \emph{unchanged}, with a 1\% 
reduction even though the overall LISA population of both variations is 
halved. A large portion of the maintenance of the number of DWDs in 
parameter variation $q3$ is contributed to by He + He binaries with higher 
GW frequencies and lower ASDs which can now survive CE interactions due to 
an initial phase of mass transfer (see discussion of this formation 
channel in Section \ref{sec:CEsep}).

The distance and chirp mass of DWDs which exhibit observable orbital 
evolution due to the emission of GWs during the LISA mission can be 
measured. This is because the chirp mass -- distance degeneracy in the 
observed strain can be broken with the observed GW frequency evolution, or 
chirp. Assuming a chirp resolution of $1/T_{\rm{obs}}^2 \sim 
8\times10^{-9}\,\rm{Hz}^2$, we select the DWDs whose chirp masses and 
distances can be measured. Figure \ref{fig:Mc_vs_dist} shows the chirp 
mass vs the luminosity distance for each DWD type in this selected 
population for the fiducial variation. The same figure for our other three 
parameter variations is shown in Appendix \ref{appendix:Mc_dist_vars}. The 
contours show the $5^{\rm{th}}$, $25^{\rm{th}}$, $50^{\rm{th}}$, 
$75^{\rm{th}}$, and $95^{\rm{th}}$ percentiles for models FZ (light blue) 
and F50 (dark blue, dashed). Despite the reduction in the height of the 
confusion foreground when considering model FZ relative to F50, we find 
that LISA is unable to differentiate between the chirp mass -- distance 
distributions of the two models.

\section{Galactic DWD Confusion Foreground Discussion}
\label{sec:model_compare}

\begin{figure}
	\includegraphics[width=0.45\textwidth]{figures/model_comp.pdf}
    \caption{The top panel shows the Galactic DWD confusion fit vs GW 
frequency for different binary evolution parameter variations (colors) 
with model F50 shown in dashed lines and model FZ shown in solid lines. 
The bottom panel shows the ratio of the number of DWDs orbiting in the 
LISA frequency band for model FZ to model F50 vs to number of DWDs 
orbiting in the LISA band for model F50 only. While both the height of the 
confusion foreground and number of LISA DWDs changes for each variation, 
the FZ models within each variation exhibit either very little change or a  
near-constant reduction by a factor of two compared with the F50 models.}
    \label{fig:model_comp}
    \script{model_comp.py}
\end{figure}

In each parameter variation, changing the binary evolution assumptions 
dramatically changes the formation and evolution of the DWD populations. 
These changes lead to large shifts in the overall number of close DWDs in 
our synthetic present-day Milky-Way-like galaxies. We find that the total 
number of DWDs with $f_{\rm{GW}}>10^{-4}\,\rm{Hz}$ increases for both 
variations $q3$ and $\alpha5$. This is because there are fewer stellar 
mergers which occur before the formation of a DWD, thus allowing more 
systems to evolve due to GW emission and orbit in the LISA frequency band 
at present. Conversely, for variation $\alpha25$, we find that the number 
of DWDs orbiting with frequencies in the LISA band is drastically reduced. 
This is because of the highly inefficient use of orbital energy to eject 
the common envelope, thus producing more stellar mergers, or closer 
binaries which are more prone to future mergers.

Interestingly, when we compare the populations of each binary fraction 
model for our variations, we find that the number of close DWDs remains 
approximately unchanged for our fiducial set of assumptions and $\alpha25$ 
variation, but is reduced by 50\% for variations $\alpha5$ and $q3$. This 
is illustrated in Figure \ref{fig:model_comp}. The top panel shows the 
confusion foreground fits for each binary evolution variation (different 
colored lines) and for each binary fraction model where the solid lines 
show FZ models and dotted lines show F50 models. The bottom panel shows 
the ratio of the number of DWDs orbiting in the LISA frequency band for 
the FZ models vs the F50 models for each variation. The number of DWDs in 
the LISA band spans over two orders of magnitude, and the ratio of FZ to 
F50, as well as the spectral shape of the confusion fit, is either 
relatively unchanged or reduced by a factor of approximately two. This 
suggests that assuming a metallicity-dependent binary fraction may change 
the size of the Galactic close DWD population by a factor of $\sim$ two 
and the strength of the Galactic DWD GW foreground for LISA depending on 
the chosen binary evolution model.

\section{Conclusions}\label{sec:conclusions}
In this study, we have investigated the effects of assuming a 
metallicity-dependent binary fraction on the formation and evolution of 
the Galactic population of DWDs with a special focus on the implications 
for LISA. Based on our synthetic Milky-Way-like galaxy catalogs of DWDs, 
we find that applying a metallicity-dependent binary fraction changes the 
formation efficiency and evolutionary history of DWD populations. However, 
when considering the close DWD populations observable by LISA, we find 
that the only distinguishing features between models which assume a 
metallicity-dependent binary fraction (model FZ) and models which assume a 
flat $50\%$ binary fraction (model F50) are the population sizes and the 
strength of the Galactic DWD GW foreground. Models which assume a 
metallicity-dependent binary fraction produce Galactic DWD populations 
that were unchanged relative to the standard model assumptions. However, 
our results are highly dependent on the chosen binary evolution model. We 
extended our study to include three binary evolution parameter variations 
to investigate the sensitivity of the DWD population changes in 
assumptions for mass transfer stability and common envelope ejection 
efficiencies. While our binary evolution parameter variations change the 
size of the LISA-observable populations dramatically, the size of the 
close Galactic DWD population and the height of the confusion foreground 
for models which assume a metallicity-dependent binary fraction appears to 
either remain unchanged \emph{or} be reduced by 50\%. This reduction 
extends to the height of the confusion foreground in the LISA data stream. 
These two results highlight the interplay between the dominating factors 
which shape the LISA DWD population: the metallicity-dependent star 
formation history of the \textbf{m12i} galaxy, the metallicity-dependent 
close binary fraction which alters the ZAMS orbital period distribution, 
and the impact of binary evolution assumptions on DWD formation. This 
comparison emphasizes the sensitivity of the Galactic population of 
LISA-detectable DWDs to the intricacies of binary evolution. Binary 
evolution assumptions like CE efficiency are undergoing continuous study 
in current literature and an exact value has yet to be identified. By 
exploring a range of possibilities in this work, we capture the variation 
in predictions of the LISA DWD population that may be observed.

An important consequence of a lower Galactic DWD confusion foreground is 
that relative to the total DWD population, more DWDs can be individually 
resolved because of the reduction in competing GW signals. While the 
number of DWDs radiating GWs in the LISA frequency band is reduced by a 
factor of two for model FZ relative to model F50 for variations $\alpha25$ 
and $q3$, the number of resolved sources is less affected with a 
population-wide reduction of $26\%$ for the former, and surprisingly only 
by 1\% for the latter. These results are far-reaching since the strength 
of the Galactic DWD confusion foreground has direct consequences on the 
detectability of all other LISA sources with small SNRs. An increase in 
resolution capability from the reduced confusion foreground can be 
extended to other Galactic binaries that LISA will observe at these 
frequencies like those involving neutron stars and stellar-origin black 
holes, as well as other more exotic GW sources like merging supermassive 
black holes, extreme mass ratio inspirals, or cosmological GW backgrounds. 
Based on our results, we suggest that studies which employ fits to the 
confusion foreground based on population synthesis results consider 
reducing the strength of the Galactic foreground PSD by a factor of two 
depending on the chosen binary evolution model.

\begin{acknowledgments}
The authors are grateful for helpful discussions with Carles Badenes, 
Christine Mazzola Daher, Kaitlin Kratter, and the Gravitational Waves and 
Astronomical Data groups at the CCA. The authors are also grateful to the 
referee for providing a thoughtful review which strengthened the 
presentation of the manuscript. S.T.\ was supported by an Undergraduate 
Student Research Award (USRA) at CITA from the Natural Sciences and 
Engineering Research Council of Canada (NSERC), Reference \# 498223. K.B.\ 
is grateful for support from the Jeffrey L. Bishop Fellowship. The 
Flatiron Institute is supported by the Simons Foundation.
\end{acknowledgments}

\section*{Data Availability}

All data and software required to reproduce our results are available 
through GitHub and Zenodo which are accessible through the icon links 
associated with our abstract and each figure.

\software{\texttt{astropy} \citep{astropy:2013, astropy:2018}; 
          \cosmic\ \citep{Breivik2020a};
          \legwork\ \citep{Wagg2021};
          \texttt{matplotlib}\ \citep{matplotlib}; 
          \texttt{numpy}\ \citep{numpy}; 
          \texttt{pandas}\ \citep{mckinney-proc-scipy-2010, 
reback2020pandas}; 
          \texttt{scipy}\ \citep{scipy}
          \texttt{seaborn}\ \citep{Waskom2021}
          }

\bibliographystyle{aasjournal}
\bibliography{bib}{}

\appendix
\section{\textbf{Formation efficiency trends}}\label{appendix:form_eff}
\textbf{As discussed in Section \ref{sec:formeff}, the formation 
efficiency of DWDs exhibitis a metallicity dependence. This is both due to 
the impact of metallicity on stellar evolution, and on the ZAMS orbital 
period distribution of model F50 vs. FZ. Each DWD type and binary 
evolution parameter variation exhibits unique trends in efficiency and in 
their evolutionary channel. In the following subsections we go through 
each DWD type in detail, first for our fiducial simulations and then for 
our variations.}

\subsection{Fiducial formation efficiency 
trends}\label{subsec:formeff_fid}

\subsubsection{\textbf{He + He}}\label{formeff_HeHe}
For He + He DWDs, the sudden drop in formation efficiency near 
$\log(Z/Z_{\odot})=-1.0$ \textbf{(see Figure \ref{fig:form_eff})} is 
generally caused by the timescale for which the initially more massive 
star in the DWD progenitor overflows it's Roche lobe. \textbf{As discussed 
in Section \ref{sec:formeff}, at} lower metallicities, donors tend to fill 
their Roche lobes while they are still on the MS and the mass transfer 
remains stable and serves to widen the binary. At higher metallicities, 
mass transfer is initiated when the donor has left the main sequence, and 
the binary enters a common envelope evolution.

The ZAMS orbital period leads to different specific \textbf{He + He} 
evolutionary channels. For short-period systems with periods below 
$10~\rm{days}$, CE leads to a stellar merger due to insufficient orbital 
energy to eject the envelope. Stellar mergers continue to dominate the 
evolutionary pathways of systems with intermediate orbital periods ($1 < 
\log_{10}(P_{\rm{orb}}/\rm{day}) < 2.5$). However, an additional growing 
number of merging systems arises with one WD component and one stellar 
companion, and \textbf{surviving} systems which don't interact at all and 
thus do not form a He + He DWD. The distinction between the various 
scenarios in this intermediate orbital period range depends on the 
combination of their ZAMS masses and orbital periods. Finally, at wider 
initial periods $\log_{10}(\rm{P}) > 2.5$, the decrease in formation 
efficiency is dominated by systems which never interact and thus do not 
form a DWD before the present day.

\subsubsection{\textbf{CO + He}}\label{formeff_COHe}
\textbf{In general, a CO + He DWD fails to form because either a stellar 
merger occurs before DWD formation, or a CO + CO or He + He DWD is formed 
instead. We discuss the specifics of these effects for three ZAMS orbital 
period regimes below. Model FZ yields higher formation efficiencies than 
F50 because CO + He DWDs prefer to form from progenitor binaries with 
orbital periods near $\log_{10}(P_{\rm{orb}}/~\rm{day})\sim2-4$, which is 
where the orbital period distribution of FZ peaks.}

For short-period CO + He DWD progenitor binaries with orbital periods 
below $\sim30~\rm{days}$, the only channel for stellar mergers before DWD 
formation is during a CE evolution. This channel is similar to the stellar 
merger channel for He + He DWDs, but due to higher-mass progenitors the CE 
evolution results in the binary component with a higher mass becoming a CO 
WD. At higher metallicities, the CO WD merges with its companion. 
\textbf{As discussed in Section \ref{sec:formeff}, h}igher-metallicity 
progenitors have a larger fraction of their mass in the convective 
envelope when compared to lower-metallicity stars of the same mass. Thus a 
CE phase with a higher-metallicity progenitor requires more orbital energy 
to eject the common envelope, causing a larger amount of orbital shrinking 
which results in a merger later in its binary evolution.

For CO + He DWD progenitor systems with intermediate orbital periods 
($1.5\leq \log_{10}(P_{\rm{orb}}/\rm{day})< 2.5$) the mechanisms which 
impact formation efficiency are complex. The dominant way CO + He DWDs 
\textbf{fail to form} is stellar mergers that occur during the second CE 
phase. \textbf{Low-metallicity systems survive this second CE phase 
whereas high-metallicity systems do not.} \textbf{Similar to the 
short-period systems, there is increased orbital shrinkage for 
high-metallicity progenitor binaries due to the deeper convective envelope 
of the CE evolution donor. This creates shorter post-CE orbital periods 
which leads to mergers during the CE second phase.}

If a lower-metallicity system forms a CO + He DWD and a higher-metallicity 
system does not, this \textbf{could also be} because the initially more 
massive binary component initiates a CE while on the giant branch instead 
of the asymptotic giant branch. This leaves behind a He WD with a stellar 
companion \textbf{instead of a CO + He}, which is the dominating scenario 
that restricts formation efficiency. There are a few edge cases where 
either a He + He DWD is formed instead, or a CO WD and a stellar companion 
is formed and there has been very nearly but not quite enough time for a 
CO + He DWD to form.

Long-period binaries with $\log_{10}(P_{\rm{orb}}/\rm{day})>2.5$ also 
display complex scenarios that hinder CO + He DWD formation. In near-equal 
contributions, our \cosmic\ simulations produce either stellar mergers or 
stable non-DWD binaries at the end of the Hubble time. At these orbital 
periods, stellar mergers always occur with a CE phase between a stellar 
companion and a CO or He WD. \textbf{Again}, the mergers occur because of 
increased CE donor envelope masses at higher metallicities. A subdominant 
channel of stable He + He DWDs can also occur when a high-metallicity 
primary overflows it's Roche lobe while still on the giant branch and thus 
forms a He WD.

\subsubsection{\textbf{CO + CO}}\label{formeff_COCO}
The decrease in the CO + CO DWD formation efficiency with increasing 
metallicity stems from different evolutionary channels which arise at the 
ZAMS orbital period boundary of $\log_{10}(P_{\rm{orb}}/\rm{day})\simeq3$. 
We find that for binaries with orbital periods below this boundary, the 
most common way that CO + CO DWDs form at lower metallicities but not at 
higher metallicities is through stellar mergers during a CE phase with a 
donor that is still on the giant branch. For lower-metallicity binaries, 
which evolve on faster timescales, the primaries enter CE evolution while 
on the asymptotic giant branch instead and the binary is able to survive. 
For binaries with orbital periods above the boundary, the vast majority of 
systems with wide initial orbits end up as stable binaries. The 
\textbf{wide} systems which don’t form a CO DWD at high metallicity do so 
because one or both of the binary components initiate a CE phase while 
still on the giant branch, thus producing a CO + He or He + He DWD.

\subsubsection{\textbf{ONe + X}}\label{formeff_ONe}
The strongest effect which hinders formations of higher-metallicity ONe + 
X DWDs is the strength of metallicity-dependent stellar winds assumed in 
our model \citep{Vink2001}. The strength of line-driven winds varies more 
strongly for the more massive ($\geq5\,M_\odot$) progenitors of ONe WDs 
relative to the other lower mass WD progenitors. At higher metallicities, 
ONe WD progenitors can lose enough mass through winds such that they don't 
ignite their CO cores and thus leave behind a CO WD. Conversely, the 
lower-metallicity progenitors retain enough mass to cause carbon ignition 
and leave behind an ONe WD.

\subsection{Parameter variations formation efficiency 
trends}\label{subsec:formeff_vars}

\textbf{As discussed in Section \ref{sec:formeff},} our simulations with 
binary evolution parameter variations \textbf{show competition between} 
the physics of binary evolution with skews introduced by the 
metallicity-dependent orbital period distribution and binary fraction. Two 
important features of our FZ and F50 models drive the differences in 
\textbf{formation efficiency for} our binary evolution parameter 
variations. First, the log-normal metallicity-dependent ZAMS orbital 
period distribution increasingly skews towards 
\textbf{$P_{\rm{orb}}\lesssim 10^4$ days} for decreasing metallicity. 
However, second, the log-uniform $P_{\rm{orb}}$ distribution of F50 
produces a higher number of the shortest-period systems 
(\textbf{$P_{\rm{orb}}\lesssim 100$ days}) than for FZ as the tail of the 
FZ distribution falls \textbf{below F50 in this regime.} \textbf{Because 
there are more ZAMS binaries with $P_{\rm{orb}} \lesssim 10^4$ days in 
F50, as the population evolves there are more likely to be binary 
interactions in the population for F50 than FZ. This leads to more mergers 
in $\alpha25$ occurring for F50 before formation of the DWD, and thus a 
lower $\alpha25$ formation efficiency overall for F50 in contrast to FZ. 
Below we summarize how the formation efficiency of each DWD type is 
impacted by the parameter variations. Across all DWD types, we find that 
there are more mergers for variation $\alpha25$ compared with the fiducial 
variation, and fewer for $\alpha5$, due to the CE efficiency's impact on 
CE phase orbital energy requirements. We also find fewer mergers for $q3$, 
because of the ability for binaries to maintain stable mass transfer, thus 
increasing their chance of survival.}

\subsubsection{\textbf{Parameter variations: He + 
He}}\label{formeff_HeHe_vars}
For binary systems leading to He + He DWDs, \textbf{as stated above} there 
are many more mergers for $\alpha25$ than for the fiducial variation. The 
lower CE efficiency leads to more DWD progenitor mergers from a failed CE 
ejection, or the CE shrinking the orbit such that it then merges on the 
second CE phase. \textbf{As discussed above, there are also more close 
binaries in the orbital period distribution of F50 than FZ, creating a 
higher propensity for binary interactions and thus mergers for 
$\alpha25$.} This is illustrated in the $\alpha25$ He + He plot of Figure 
\ref{fig:form_eff}. Similar to the fiducial parameter variation, a binary 
that becomes a He + He DWD at lower metallicities would merge before 
becoming one at higher metallicities because of the longer evolution 
timescales. These later interactions lead to closer post-CE separations 
and thus more mergers. The dip at [Fe/H] $\simeq -1.5$ \textbf{(see Figure 
\ref{fig:form_eff})} stems from mergers before the DWD forms, again 
similar to the fiducial parameter variation. The peak that occurs in 
$\alpha25$ at $\sim 10\%\,Z_\odot$ arises from binaries evolving faster, 
and thus interacting earlier, as metallicity increases. This means that 
equivalent binaries that would have merged or formed a CO + He DWD instead 
at lower metallicities now survive to produce a He + He DWD. At higher 
metallicities, there are also a higher proportion of wide ZAMS binaries, 
which evolve off the Main Sequence too slowly to form a He + He DWD, 
\textbf{adding to} the efficiency drop.

For $\alpha5$, there are significantly fewer DWD progenitor mergers since 
binaries are more likely to survive a CE phase due to the higher CE 
ejection efficiency. Systems that would merge for $\alpha25$ instead 
survive, shrinking to short orbits through CE interactions to reside in 
the LISA band at present. The number of DWDs then makes up a higher 
fraction of the total simulated binary population, thus we find in general 
higher formation efficiencies than for the fiducial parameter variation 
which sets $\alpha=1$. This effect is even more pronounced in comparison 
with the $\alpha25$ variation which sets $\alpha=0.25$.

Beyond the effects described above, the efficiency drop across metallicity 
is similar to the fiducial parameter variation in that it depends on the 
metallicity-dependent ZAMS orbital period distribution emphasizing 
short-period systems at low metallicities, and on the change of formation 
channel for different period ranges across metallicity. We find the 
variation $q3$ yields the highest formation efficiency. Only the 
shortest-period binaries do not form a DWD\textbf{, which are lost} 
through mergers at higher metallicities. At low metallicities, even 
binaries with ZAMS orbital periods $P_{\rm{orb}}\lesssim 100$ days form He 
+ He DWDs, with binaries in model FZ reaching ZAMS orbital periods less 
than a day in length. This leads to a higher formation efficiency from F50 
over FZ for parameter variation $q3$ because of the higher number of 
shortest-period ZAMS binaries in the log-uniform distribution of F50.

\subsubsection{\textbf{Parameter variations: CO + 
He}}\label{formeff_COHe_vars}
For CO + He binaries, there are again many more mergers for parameter 
variation $\alpha25$ relative to the fiducial case. This is the dominant 
\textbf{reason} that progenitors do not evolve to produce DWDs at higher 
metallicities. At low metallicities, a number of progenitor binaries form 
CO + CO DWDs instead of He + CO DWDs. This is especially the case near 
$\log_{10}(Z/Z_\odot)\sim -2$, where RLO occurs late enough in a binary's 
evolution that mass transfer remains stable and CO + CO's are formed 
rather than CO + He, hence producing a dip in the formation efficiency 
\textbf{(see Figure \ref{fig:form_eff})}.

Because parameter variation $\alpha5$ \textbf{yields fewer} mergers, the 
formation efficiency is mostly governed by the metallicity-dependent ZAMS 
orbital period distribution. At higher metallicities, the majority of 
cases in model FZ where progenitors do not produce a CO + He DWD are from 
progenitor mergers, or from He + He DWDs forming instead of CO + He DWDs. 
At higher metallicities, the ZAMS orbital period distribution of FZ also 
drops below that of F50 and extends out to orbital periods beyond what 
would form a CO + He. This causes the formation efficiencies for FZ and 
F50 to cross at log$_{10}(Z/Z_\odot)\sim-0.5$. This is also apparent for 
the $q3$ variation, and to a lesser extent in the fiducial case. In $q3$, 
for any given orbital period we find that lower-mass DWDs originate from 
lower metallicities, thus leading to an approximately constant formation 
efficiency.

\subsubsection{\textbf{Parameter variations: CO + 
CO}}\label{formeff_CO_vars}
For CO + CO, across all variations, very close ZAMS binaries 
($P_{\rm{orb}}\sim\mathcal{O}(100$ days)) at high metallicities form a CO 
+ He or He + He DWD instead of a CO + CO because the binary interacts 
before core Helium burning of one or both components. There are again many 
more mergers in $\alpha25$, and fewer mergers for $\alpha5$. Due to the 
high CE efficiency of $\alpha5$, systems with shorter ZAMS orbital periods 
survive to become CO + CO DWDs rather than merging or becoming CO + He or 
He + He DWDs. There are fewer mergers overall for $q3$ than all the 
$\alpha$ variations because the first RLO occurrence leads to stable mass 
transfer from the primary to the secondary rather than multiple CE phases. 
For both the $\alpha5$ and $q3$ variations, the majority of CO + CO DWDs 
originate in wide enough orbits that they do not interact before DWD 
formation. This leads to an approximately constant formation efficiency 
across metallicity in both cases.

\subsubsection{\textbf{Parameter variations: ONe + 
X}}\label{formeff_ONe_vars}
For model FZ, the ONe + X formation efficiencies follow similar trends 
aross each parameter variation. Their formation scenarios are similar to 
those described \textbf{in Section \ref{sec:formeff}} for ONe + X DWDs in 
the fiducial binary evolution parameter set. The main difference between 
each parameter variation is only in the number of sources produced since 
the number of stellar mergers that occur before DWD formation increases 
with decreasing common envelope efficiency. In model F50, the formation 
efficiency is decreased significantly for parameter variation $\alpha25$ 
because the log-uniform orbital period distribution produces many ZAMS 
binaries with very short orbital periods which lead to stellar mergers 
during the CE phases.

\section{Amplitude Spectral Densities for Binary Evolution Parameter 
Variations}\label{appendix:ASD_vars}

Figure \ref{fig:LISA_SNR_vars} shows the ASD vs GW frequency of the 
resolved systems with SNR $> 7$ for each DWD type, where the rows 
alternate between models F50 and FZ for the $\alpha25$, $\alpha5$, and 
$q3$ parameter variations. The LISA sensitivity curve is also shown in 
black, which includes the modeled confusion foreground \textbf{(see 
Section \ref{sec:LISA_met} for details on fitting the foreground) for the 
given variation}. The entire population is shown as grey scatter points. 
As discussed in Section \ref{sec:LISA_met}, apart from each DWD type 
having a different abundance of resolved systems, the population-wide 
characteristics remain unchanged between the two binary fraction models.

The height of the confusion foreground and the number of DWDs radiating 
GWs with $f_{\rm{GW}} > 10^{-4}\,\rm{Hz}$ is reduced by a factor of 
\textbf{two} for model FZ relative to model F50 for variations $\alpha5$ 
and $q3$, however the number of resolved\textbf{, i.e. observable,} 
sources is not reduced to an equal degree. The number of observable 
systems with SNR $>7$ increases by 3\% for $\alpha25$\textbf{, mimicking 
the LISA-band population remaining relatively unchanged}. \textbf{However, 
}the number of \textbf{observable} DWDs is reduced by only 26\% for 
$\alpha5$, and \textbf{by only 1\% for $q3$}, even though the overall LISA 
population of both variations is halved. \textbf{Maintaining} the number 
of DWDs in parameter variation $q3$ is contributed to \textbf{largely} by 
He + He binaries with higher GW frequencies and lower ASDs which can now 
survive CE interactions due to an initial phase of mass transfer (see 
discussion of this formation channel in Section \ref{sec:CEsep}).

\begin{figure*}[h]
    \centering

\includegraphics[width=0.8\textwidth]{figures/LISA_SNR_alpha25.pdf}
	\includegraphics[width=0.8\textwidth]{figures/LISA_SNR_alpha5.pdf}
	\includegraphics[width=0.8\textwidth]{figures/LISA_SNR_q3.pdf}
    \caption{The ASD vs GW frequency for DWDs resolved with SNR $> 7$ for 
each DWD type and parameter variation. The rows alternate between model 
F50 and from model FZ. In each panel, the LISA sensitivity curve, 
including the confusion foreground for each model, is shown in black and 
the total population for each model is shown in grey. Again, we find that 
each model qualitatively exhibits similar characteristics and that the 
only change is in the yield of resolved DWDs for each type based on the 
strength of the confusion foreground.}
    \label{fig:LISA_SNR_vars}
     \script{LISA_SNR.py}

\end{figure*}

\section{Chirp Mass -- Distance Distributions for Binary Evolution 
Parameter Variations}\label{appendix:Mc_dist_vars}

Figure~\ref{fig:Mc_vs_dist_vars} shows the chirp mass vs distance 
distribution for the $\alpha25$, $\alpha5$, and $q3$ parameter variations. 
For $\alpha25$ we plot scatter plots for He + He and ONe + X due to low 
statistics; there are too few binaries in this selected population to 
produce reliable kernel density estimates. The only population with a 
significant change between \textbf{binary fraction models} is for ONe + X 
DWDs\textbf{, variation} $\alpha5$. Here, we find a skew to lower 
\textbf{chirp} masses because a higher fraction of the binaries are ONe + 
He for FZ \textbf{compared to} F50: a Helium companion makes up 44\% of 
this selected population for FZ, compared with 10\% for F50. This leads to 
lower average chirp masses for the ONe + X distribution. Furthermore, 
$\sim8$\% of the F50 measurable systems are ONe + ONe binaries, versus 0\% 
for FZ. Because ONe WDs have larger masses, this also creates the 
high-mass concentration we \textbf{find} in the F50 contours.
\begin{figure*}[h]
    \centering

\includegraphics[width=0.8\textwidth]{figures/Mc_vs_dist_alpha25.pdf}

\includegraphics[width=0.8\textwidth]{figures/Mc_vs_dist_alpha5.pdf}
	\includegraphics[width=0.8\textwidth]{figures/Mc_vs_dist_q3.pdf}
    \caption{Chirp mass -- distance distributions for our other three 
binary evolution parameter variations. We plot the $\alpha25$ populations 
of He + He and ONe + X DWDs as scatter points since there are too few 
binaries to produce meaningful density distributions. Again, we have 
plotted each DWD type's population that exhibit observable GW frequency 
evolution and have $\rm{SNR}>7$. Contours show the 5$^{\rm{th}}$, 
25$^{\rm{th}}$, 50$^{\rm{th}}$, 75$^{\rm{th}}$ and 95$^{\rm{th}}$ 
percentiles. Most populations remain unchanged between binary fraction 
models except for ONe + X DWDs in $\alpha5$, due to the number of He vs. 
CO vs. ONe companions between each model.}
    \label{fig:Mc_vs_dist_vars}
    \script{Mc_vs_dist.py}
\end{figure*}
\end{document}

\end{document}